---
title: "Bishop_Lab2"
author: "Erica Bishop"
date: "2023-01-18"
output: pdf_document
echo: TRUE
editor_options: 
  markdown: 
    wrap: 72
---

Today we will be continuing the pumpkin case study from last week. We will be using the data that you cleaned and split last time (pumpkins_train) and will be comparing our results today to those you have already obtained, so open and run your Lab 1 .Rmd as a first step so those objects are available in your Environment (unless you created an R Project last time, in which case, kudos to you!).

```{r}
#run lab 1 code into environment

sourceDir <- "/Users/ericabishop/Documents/MEDSwinter/EDS232-ml/labs/Lab1.Rmd"
library(knitr)
source(knitr::purl(sourceDir, quiet=TRUE))


```

Once you have done that, we'll start today's lab by specifying a recipe for a polynomial model. First we specify a recipe that identifies our variables and data, converts package to a numerical form, and then add a polynomial effect with step_poly()

```{r}
# Specify a recipe
poly_pumpkins_recipe <-
  recipe(price ~ package, data = pumpkins_train) %>%
  step_integer(all_predictors(), zero_based = TRUE) %>% 
  step_poly(all_predictors(), degree = 4)


```

How did that work? Choose another value for degree if you need to. Later we will learn about model tuning that will let us do things like find the optimal value for degree. For now, we'd like to have a flexible model, so find the highest value for degree that is consistent with our data.

Polynomial regression is still linear regression, so our model specification looks similar to before.

```{r}
# Create a model specification called poly_spec
poly_spec <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")
```

Question 1: Now take the recipe and model specification that just created and bundle them into a workflow called poly_df.

```{r}
# Bundle recipe and model spec into a workflow
poly_wf <- workflow() |> 
  add_recipe(poly_pumpkins_recipe) |> 
  add_model(poly_spec)

```

Question 2: fit a model to the pumpkins_train data using your workflow and assign it to poly_wf_fit

```{r}
# Create a model
poly_wf_fit <- poly_wf |> 
  fit(data = pumpkins_train)
  
```

```{r}
# Print learned model coefficients
poly_wf_fit
```

```{r}
# Make price predictions on test data
poly_results <- poly_wf_fit %>% predict(new_data = pumpkins_test) %>% 
  bind_cols(pumpkins_test %>% select(c(package, price))) %>% 
  relocate(.pred, .after = last_col())

# Print the results
poly_results %>% 
  slice_head(n = 10)
```

Now let's evaluate how the model performed on the test_set using yardstick::metrics().

```{r}
poly_metrics <- metrics(data = poly_results, truth = price, estimate = .pred)

print(poly_metrics)

#compare to the results from the linear model
lm_metrics <- metrics(data = lm_results, truth = price, estimate = .pred)
print(lm_metrics)
```

Question 3: How do the performance metrics differ between the linear model from last week and the polynomial model we fit today? Which model performs better on predicting the price of different packages of pumpkins?

**The polynomial model fits the data better, with a much higher R squared value (0.89 compared to 0.49). The root mean square error is also much smaller for the polynomial model, meaning there is less variance in the errors. The mean absolute errors (MAE) is also smaller for the polynomial model than the linear model.**

Let's visualize our model results. First prep the results by binding the encoded package variable to them.

```{r}
# Bind encoded package column to the results
poly_results <- poly_results %>% 
  bind_cols(package_encode %>% 
              rename(package_integer = package)) %>% 
  relocate(package_integer, .after = package)


# Print new results data frame
poly_results %>% 
  slice_head(n = 5)
```

OK, now let's take a look!

Question 4: Create a scatter plot that takes the poly_results and plots package vs. price. Then draw a line showing our model's predicted values (.pred). Hint: you'll need separate geoms for the data points and the prediction line.

```{r}
# Make a scatter plot
poly_results |> ggplot(
       aes(x = package_integer,
           y = price)) +
  geom_point() +
  geom_line(aes(y = .pred))

```

You can see that a curved line fits your data much better.

Question 5: Now make a smoother line by using geom_smooth instead of geom_line and passing it a polynomial formula like this:
geom_smooth(method = lm, formula = y \~ poly(x, degree = 3), color = "midnightblue", size = 1.2, se = FALSE)

```{r}
# Make a smoother scatter plot 

poly_results |> ggplot(
       aes(x = package_integer,
           y = price)) +
  geom_point() +
  geom_smooth(method = lm, formula = y ~ poly(x, degree = 3),
              color = "midnightblue",
              size = 1.2,
              se = FALSE)
  
```

OK, now it's your turn to go through the process one more time.

Additional assignment components : 

6. Choose a new predictor variable (anything not involving package type) in this dataset. 

**Predictor variable: variety**

7. Determine its correlation with the outcome variable (price). (Remember we calculated a correlation matrix last week) 

```{r}

price_var <- cor(baked_pumpkins$price, baked_pumpkins$variety)

print(paste("There is a", round(price_var, 3), "correlation between pumpkin price and variety."))
```

8. Create and test a model for your new predictor:
- Create a recipe 
```{r}
set.seed(321) #sets starting point for random number generation (to randomly split data)
# Split the data into training and test sets
pumpkins_split2 <- new_pumpkins %>% 
  initial_split(prop = 0.8)

# Extract training and test data
ptrain <- training(pumpkins_split2)
ptest <- testing(pumpkins_split2)


#Using a linear recipe to make the encode column work
#there are only two varieties (miniature and pie - so a linear works better here than polynomial)
#or at least I think that's why I can't get the plotting to work with a polynomial recipe
lin_pumvarrecipe <- recipe(price ~ variety,
                           data = ptrain) |> 
  step_integer(all_predictors(), zero_based = TRUE)


```
- Build a model specification (linear or polynomial) 

```{r}
# Create a linear model specification
mod_spec <- linear_reg() |>  
  set_engine("lm") |>  
  set_mode("regression")

```

- Bundle the recipe and model specification into a workflow 

```{r}

mod_wf <- workflow() %>% 
  add_recipe(lin_pumvarrecipe) %>% 
  add_model(mod_spec)

```

- Create a model by fitting the workflow 
```{r}

# Train the model
mod_wf_fit <- mod_wf |> 
  fit(data = ptrain)

# Print the model coefficients learned 
print(mod_wf_fit)


```
- Make predictions with the test data

```{r}
# Make price predictions on test data
mod_results <- mod_wf_fit %>% predict(new_data = ptest) %>% 
  bind_cols(ptest %>% select(c(variety, price))) %>% 
  relocate(.pred, .after = last_col())

head(mod_results)

```
- Evaluate model performance on the test data 

```{r}

# Evaluate performance of linear regression
metrics(data = mod_results,
        truth = price,
        estimate = .pred)

```

- Create a visualization of model performance

```{r}
#need to encode variety column in order to visualize

#Encode variety column with pumvar_recipe
variety_encode <- lin_pumvarrecipe %>%  #WHY won't this step work with a polynomial recipe? - is it because there are only two varieties?
  prep() %>% 
  bake(new_data = ptest) %>% 
  select(variety)

# Bind encoded package column to the results
 mod_plot <- mod_results %>%
 bind_cols(variety_encode %>%
               rename(variety_integer = variety)) %>%
  relocate(variety_integer, .after = variety)


# Print new results data frame
print(head(mod_plot))


# Make a scatter plot
mod_plot %>%
  ggplot(aes(x = variety_integer,
            y = price)) +
   geom_point(size = 1.6) +
   geom_line(aes(y = .pred), 
             color = "orange", 
             size = 1.2) +
   xlab("pumpkin variety") +
  ylab("price")



```


Lab 2 due 1/24 at 11:59 PM
