---
title: "Lab 4"
author: "Erica Bishop"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(skimr)
library(tidymodels)
library(caret)
library(corrplot)
library(gt)
library(broom)
library(patchwork)
```

## Lab 4: Fire and Tree Mortality

The database we'll be working with today includes 36066 observations of individual trees involved in prescribed fires and wildfires occurring over 35 years, from 1981 to 2016. It is a subset of a larger fire and tree mortality database from the US Forest Service (see data description for the full database here: [link](https://www.nature.com/articles/s41597-020-0522-7#Sec10)). Our goal today is to predict the likelihood of tree mortality after a fire.

### Data Exploration

Outcome variable: *yr1status* = tree status (0=alive, 1=dead) assessed one year post-fire.

Predictors: *YrFireName, Species, Genus_species, DBH_cm, CVS_percent, BCHM_m, BTL* (Information on these variables available in the database metadata ([link](https://www.fs.usda.gov/rds/archive/products/RDS-2020-0001-2/_metadata_RDS-2020-0001-2.html))).

```{r}
trees_dat<- read_csv(file = "https://raw.githubusercontent.com/MaRo406/eds-232-machine-learning/main/data/trees-dat.csv")
```

> Question 1: Recode all the predictors to a zero_based integer form

```{r}

trees_encoded <- recipe(yr1status ~ ., data = trees_dat) |> 
  step_integer(all_predictors(), zero_based = TRUE) |> 
  prep(trees_dat) |> 
  bake(trees_dat) |> 
  select(-"...1") #drop index column

```

### Data Splitting

> Question 2: Create trees_training (70%) and trees_test (30%) splits for the modeling

```{r}

set.seed(123)  # for reproducibility (random sample)
trees_split <- initial_split(trees_encoded, prop = .70)

#create training set
trees_train <- training(trees_split)

#create testing set
trees_test  <- testing(trees_split)

```

> Question 3: How many observations are we using for training with this split?

```{r}
print(paste("We are using", nrow(trees_train), "observations for training and ", nrow(trees_test), "for testing."))

```

### Simple Logistic Regression 

Let's start our modeling effort with some simple models: one predictor and one outcome each.

> Question 4: Choose the three predictors that most highly correlate with our outcome variable for further investigation.

```{r}
#create a correlation matrix to easily identify variables most highly correlated
cormatrix <- cor(trees_encoded)

# Make a correlation plot between the variables
corrplot(cormatrix, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45, addCoef.col = "black", cl.pos = "n", order = "original")

```
The three variables most highly correlated with tree survival are:
CVS_percent: 0.68
BCHM_m: 0.42
DBH_cm: -0.32

**CVS_percent** is the percent of the pre-fire crown volume that was scorched or consumed by fire (values 0 to 100). If measured, this is the CVS from field measurements. Otherwise it is the calculated CVS from crown length measurement, where CVS=100[(CLS(2CL_pre - CLS))/CL_pre2].
**BCHM_m** is maximum bark char (also called bole char, bole scorch in other publications) vertical height from ground on a tree bole, rounded to nearest 0.01 m (m=meters).
**DBH_cm** is the diameter at breast height rounded to nearest 0.1 cm (cm = centimeters).

> Question 5: Use glm() to fit three simple logistic regression models, one for each of the predictors you identified.

```{r}

lm_cvs <- glm(yr1status ~ CVS_percent, family = "binomial", data = trees_train)

lm_bchm <- glm(yr1status ~ BCHM_m, family = "binomial", data = trees_train)
  
lm_dbh <- glm(yr1status ~ DBH_cm, family = "binomial", data = trees_train)

```


### Interpret the Coefficients 

We aren't always interested in or able to interpret the model coefficients in a machine learning task. Often predictive accuracy is all we care about.

> Question 6: That said, take a stab at interpreting our model coefficients now.

**Interpreting the crown volume model:**

```{r}

lm_cvs |> 
  broom::tidy() |> 
  gt()

exp(coef(lm_cvs)) #exponentiate coeficients for interpretation 
```

For the model that regresses tree survival on the percent of tree crown volume burned provides the generalized formula: 
$$
\operatorname{logit}(\hat p)=\log \left(\frac{\hat p}{1-\hat p}\right)= -6.61 + 0.0762 x
$$
The coeffi

**Interpreting the bark char height model:**

```{r}
lm_bchm |> 
  broom::tidy() |> 
  gt()

exp(coef(lm_bchm)) 
```

**Interpreting the breast-height diameter**

```{r}
lm_dbh |> 
  broom::tidy() |> 
  gt()

exp(coef(lm_dbh))
```

> Question 7: Now let's visualize the results from these models. Plot the fit to the training data of each model.

```{r plot}

cvs_plot <- ggplot(data = trees_train, 
                   aes(x = CVS_percent,
                       y = yr1status)) + 
  geom_point() +
  stat_smooth(method="glm",  se=TRUE,
              method.args = list(family=binomial)) +
  labs(title = "Tree survival predicted by crown volume burned, bark char height, and diameter",
       x = "Percent of crown volume burned",
       y = "One year survival status") +
  theme_minimal()

bchm_plot <- ggplot(data = trees_train, 
                   aes(x = BCHM_m,
                       y = yr1status)) + 
  geom_point() +
  stat_smooth(method="glm",  se=TRUE,
              method.args = list(family=binomial)) +
  labs(
       x = "Maximum height of bark char (m)",
       y = "One year survival status") +
  theme_minimal()

dbh_plot <- ggplot(data = trees_train, 
                   aes(x = DBH_cm,
                       y = yr1status)) + 
  geom_point() +
  stat_smooth(method="glm",  se=TRUE,
              method.args = list(family=binomial)) +
  labs(
       x = "Diameter at breast height (cm)",
       y = "One year survival status") +
  theme_minimal()

cvs_plot + bchm_plot + dbh_plot + patchwork::plot_layout(ncol = )

```


### Multiple Logistic Regression

Let's not limit ourselves to a single-predictor model. More predictors might lead to better model performance.

> Question 8: Use glm() to fit a multiple logistic regression called "logistic_full", with all three of the predictors included. Which of these are significant in the resulting model?

```{r}

logistic_full <- glm(
  yr1status ~ CVS_percent + BCHM_m + DBH_cm,
  family = "binomial",
  data = trees_train
)

tidy(logistic_full)
logistic_full |>
  broom::tidy() |>
  gt()

```


### Estimate Model Accuracy

Now we want to estimate our model's generalizability using resampling.

> Question 9: Use cross validation to assess model accuracy. Use caret::train() to fit four 10-fold cross-validated models (cv_model1, cv_model2, cv_model3, cv_model4) that correspond to each of the four models we've fit so far: three simple logistic regression models corresponding to each of the three key predictors (CVS_percent, DBH_cm, BCHM_m) and a multiple logistic regression model that combines all three predictors.

```{r}

```



>
> Question 10: Use caret::resamples() to extract then compare the classification accuracy for each model. (Hint: resamples() wont give you what you need unless you convert the outcome variable to factor form). Which model has the highest accuracy?

Let's move forward with this single most accurate model.

> Question 11: Compute the confusion matrix and overall fraction of correct predictions by the model.
>
> Question 12: Explain what the confusion matrix is telling you about the types of mistakes made by logistic regression.
>
> Question 13: What is the overall accuracy of the model? How is this calculated?

### Test Final Model

Alright, now we'll take our most accurate model and make predictions on some unseen data (the test data).

> Question 14: Now that we have identified our best model, evaluate it by running a prediction on the test data, trees_test.
>
> Question 15: How does the accuracy of this final model on the test data compare to its cross validation accuracy? Do you find this to be surprising? Why or why not?