---
title: "Lab 5 - part 1"
author: "Erica Bishop (and Jillian Allison)"
date: "2023-02-07"
output: pdf_document
---

This week's lab is a musical lab. You'll be requesting data from the Spotify API and using it to build k-nearest neighbor and decision tree models.

In order to use the Spotify you must have a Spotify account. If you don't have one, sign up for a free one here: <https://www.spotify.com/us/signup>

Once you have an account, go to Spotify for developers (<https://developer.spotify.com/>) and log in. Click the green "Create a Client ID" button to fill out the form to create an app create an app so you can access the API.

On your developer dashboard page, click on the new app you just created. On the app's dashboard page you will find your Client ID just under the header name of your app. Click "Show Client Secret" to access your secondary Client ID. When you do this you'll be issued a Spotify client ID and client secret key.

You have two options for completing this lab.

**Option 1**: **Classify by users**. Build models that predict whether a given song will be in your collection vs. a partner in class. This requires that you were already a Spotify user so you have enough data to work with. You will download your data from the Spotify API and then exchange with another member of class.

**Option 2**: **Classify by genres**. Build models that predict which genre a song belongs to. This will use a pre-existing Spotify dataset available from Kaggle.com (<https://www.kaggle.com/datasets/mrmorj/dataset-of-songs-in-spotify>)

```{r}
library(spotifyr) #API interaction
library(tidyverse)
library(tidymodels)
library(here)
library(tune)

```

Client ID and Client Secret are required to create and access token that is required to interact with the API. You can set them as system values so we don't have to do provide them each time.

Sys.setenv(SPOTIFY_CLIENT_ID = '2e066a0ebf86473990f30eee626c4ab9') Sys.setenv(SPOTIFY_CLIENT_SECRET = 'bb6735cb728a42089e0cc02680c97844')

access_token \<- get_spotify_access_token() #takes ID and SECRET, sends to Spotify and receives an access token

> *This may result in an error:*
>
> INVALID_CLIENT: Invalid redirect URI
>
> *This can be resolved by editing the callback settings on your app. Go to your app and click "Edit Settings". Under redirect URLs paste this: <http://localhost:1410/> and click save at the bottom.*

```{r}
Sys.setenv(SPOTIFY_CLIENT_ID = '5cff8a26d7c940f5b487a4e97b821719')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '851d5f1aa804476d8d1ba291a3183a2e')

access_token <- get_spotify_access_token(client_id = Sys.getenv("SPOTIFY_CLIENT_ID"),
                                         client_secret = Sys.getenv("SPOTIFY_CLIENT_SECRET"))

```

**Option 1: Data Preparation**

You can use get_my_saved_tracks() to request all your liked tracks. It would be good if you had at least 150-200 liked tracks so the model has enough data to work with. If you don't have enough liked tracks, you can instead use get_my_recently_played(), and in that case grab at least 500 recently played tracks if you can.

The Spotify API returns a dataframe of tracks and associated attributes. However, it will only return up to 50 (or 20) tracks at a time, so you will have to make multiple requests. Use a function to combine all your requests in one call.

Once you have your tracks, familiarize yourself with this initial dataframe. You'll need to request some additional information for the analysis. If you give the API a list of track IDs using get_track_audio_features(), it will return an audio features dataframe of all the tracks and some attributes of them.

These track audio features are the predictors we are interested in, but this dataframe doesn't have the actual names of the tracks. Append the 'track.name' column from your favorite tracks database.

Find a class mate whose data you would like to use. Add your partner's data to your dataset. Create a new column that will contain the outcome variable that you will try to predict. This variable should contain two values that represent if the track came from your data set or your partner's.

**Option 2: Data preparation**

Download the Spotify dataset from <https://www.kaggle.com/datasets/mrmorj/dataset-of-songs-in-spotify>

Inspect the data. Choose two genres you'd like to use for the classification task. Filter down the data to include only the tracks of that genre.

```{r}
#Get saved track list
offsets_list <- c(seq(0, 2900, 50)) #create list to get saved songs

saved_tracks_eb <- lapply(X = offsets_list,
                           FUN = get_my_saved_tracks, 
                           limit = 50
                           ) |> #produce list of dfs for each call of 50
  bind_rows() #bind lists into one df

#Get saved track attributes

track_id_list <- saved_tracks_eb$track.id #create list of track ids

track_attributes_eb <- lapply(X = track_id_list,
                              FUN = get_track_audio_features) |> 
  bind_rows()


#now append track list name to the track attributes df

saved_track_names <- saved_tracks_eb |> 
  select(track.name, track.id) |> #select out just track name and id
  rename(id = track.id) #rename id column to join

eb_tracks_df <- left_join(track_attributes_eb, saved_track_names, by = "id") #join df

#add user column
eb_tracks_df <- eb_tracks_df |> 
  add_column(user = "Erica")


#save dataframe as a csv to send to Jillian
write_csv(eb_tracks_df, here("eb_track_attributes.csv"))

#read in Jillian's data
ja_tracks_df <- read_csv(here("jillian_spotify_data.csv")) |> 
  select(-track_id) |> #drop track_id column (repeats id column)
  rename(track.name = track_name)#renametrack_name to match


```

##Data Exploration (both options)

Let's take a look at your data. Do some exploratory summary stats and visualization.

For example: What are the most danceable tracks in your dataset? What are some differences in the data between users (Option 1) or genres (Option 2)?

```{r}

#skim both datasets
skimr::skim(eb_tracks_df)
skimr::skim(ja_tracks_df)

#take a closer look at some variables
names(eb_tracks_df)
names(ja_tracks_df)

#do we have any of the same tracks saved??
eb_ja_shared_faves <- semi_join(eb_tracks_df, ja_tracks_df, by = "id") #use semi_join to get df of tracks in common
#looks like a lot of beyonce, frank ocean, and lady gaga, but let's take a closer look!

#what are the characteristics of our shared tracks?
eb_ja_summary <- eb_ja_shared_faves |> 
  select(danceability, energy, key, valence, loudness, speechiness, acousticness) |> 
  summarise_all(mean) |> #calculate the average of some variables of interest
  add_column(user = "both <3") #update user column to show both of us

#compare the characteristics of shared tracks to just mine
eb_summary <- eb_tracks_df |> 
  select(danceability, energy, key, valence, loudness, speechiness, acousticness) |> 
  summarise_all(mean) |> 
  add_column(user = "Erica")

#calculate same summary table for jillian
ja_summary <- ja_tracks_df |> 
  select(danceability, energy, key, valence, loudness, speechiness, acousticness) |> 
  summarise_all(mean) |> 
  add_column(user = "Jillian")

#combine the summary stats tables into one df 
summary_df <- rbind(eb_ja_summary, eb_summary, ja_summary)

```
```{r}

#combine both data sets

eb_ja_tracks <- full_join(eb_tracks_df,
                          ja_tracks_df,
                          by = c("danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "type", "id", "uri", "track_href", "analysis_url", "duration_ms", "time_signature", "track.name")) |> 
  mutate(user = case_when( 
    user.y == "Jillian" & user.x == "Erica" ~ "Both", #relabel songs in both user lists with case_when
    user.y == "Jillian" & is.na(user.x) ~ "Jillian",
    TRUE ~ "Erica"
  )) |>  #drop individual user columns and a few other less useful predictors
  select(-c(user.x, user.y, uri, track_href, analysis_url, type))

#save output df as a csv (in case api causes issues again)
write_csv(eb_ja_tracks, here("eb_ja_tracks.csv"))

```


```{r}
### compare the spread of some key variables 

#compare danceability
dance_plot <- ggplot(data = eb_ja_tracks,
       aes(y = danceability,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Daceability")

#compare energy
energy_plot <- ggplot(data = eb_ja_tracks,
       aes(y = energy,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Energy")

#compare key
key_plot <- ggplot(data = eb_ja_tracks,
       aes(y = key,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Key")

#compare valence
valence_plot <- ggplot(data = eb_ja_tracks,
       aes(y = valence,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Valence")

#compare tempo
tempo_plot <- ggplot(data = eb_ja_tracks,
       aes(y = tempo,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Tempo")

#compare mode

mode_plot <- ggplot(data = eb_ja_tracks,
       aes(y = loudness,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Loudness")

dance_plot + energy_plot + key_plot + valence_plot + tempo_plot + mode_plot

```
Based on how similar these variables are across Jillian and me, it could be very tough for the models to correctly classify tracks - but we'll see!

### **Modeling**
#Now with an added random forest component and some clarification your task.

Create competing models that predict whether a track belongs to:

**Option 1.  you or your partner's collection**

Option 2.  genre 1 or genre 2

Create three final candidate models:

1. k-nearest neighbor 
2. decision tree
3. bagged tree
    - bag_tree()
    - Use the "times =" argument when setting the engine during model specification to specify the number of         trees. The rule of thumb is that 50-500 trees is usually sufficient.  The bottom of that range should          be sufficient here.  
4. random forest
    - rand_forest()
    - m_try() is the new hyperparameter of interest for this type of model.  Make sure to include it in your         tuning process

Go through the modeling process for each model:

Preprocessing. You can use the same recipe for all the models you create. 

Resampling. Make sure to use appropriate resampling to select the best version created by each algorithm.

Tuning. Find the best values for each hyperparameter (within a reasonable range).

Compare the performance of the four final models you have created.  

Use appropriate performance evaluation metric(s) for this classification task. A table would be a good way to display your comparison.  Use at least one visualization illustrating your model results.

## Pre-Processing
Splitting the data and encoding - this pre-processed data will be used in each of the four models below.

```{r preprocessing}
#remove unique identifiers from dataset (track id and song name)
tracks_df <- eb_ja_tracks |> 
  select(-c(id, track.name)) |> 
  mutate(user = as.factor(user)) #make outcome variable a factor

#split the data FIRST
set.seed(123) #set seed

tracks_split <- initial_split(tracks_df) #split
tracks_train <- training(tracks_split) #training dataset
tracks_test <- testing(tracks_split) #testing dataset

#Preprocess the data - encoding with the recipe
tracks_recipe <- recipe(user ~., data = tracks_train) |> 
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) |> #one hot encoding of nominal variables (- user)
  step_normalize(all_numeric(), -all_outcomes()) |> #normalize scale of numeric variables 
  prep()

#bake training data
baked_tracks <- bake(tracks_recipe, tracks_train)

#bake testing data too - for later ;)
baked_test_tracks <- bake(tracks_recipe, tracks_test)

```

## KNN Model

```{r knn_model}

#set engine and specifications for KNN model
knn_spec <- nearest_neighbor() |> 
  set_engine("kknn") |> #specify how with engine
  set_mode("classification") #specify classificaton

#fit a simple model without tuning - what happens?
#what happens when you just fit this model?
knn_fit <- knn_spec %>% 
  fit(user ~. , data = tracks_train)

knn_fit

```

```{r knn_resample_tune}

#resample
## This resampling can be used for all of the models below!
set.seed(123)
cv_folds <-tracks_train |> 
  vfold_cv() #10 folds is default

#create a workflow to tune
knn_workflow <- workflow() |> 
  add_model(knn_spec) |> 
  add_recipe(tracks_recipe)


#fit resamples and tuning grid to workflow
knn_rs_wf <- workflow() |> 
  add_model(knn_spec) |> #add specification defined above
  add_recipe(tracks_recipe) |> #add recipe
  fit_resamples(
    resamples = cv_folds, #fit to resampled data
    control = control_resamples(save_pred = TRUE)
  ) 



#|>
  # tune_grid( 
  #   cv_folds,
  #   grid = data.frame(neighbors = c(1, 5, seq(10, 100, 10))) #test k values up to 100 by 10s
  # )
 
#check out the wf
knn_final_wf

```

```{r}
knn_rs_wf |> collect_metrics()
```


```{r knn_metrics}

# Check the performance with collect_metrics()
fit_knn_cv %>% collect_metrics()

#table of metrics

#at least one visual of results

```

## Decision Tree


```{r dtree_model}

#starting with a fixed specification model
tree_spec_fixed <- decision_tree(
  cost_complexity = 0.1,
  tree_depth = 4,
  min_n = 11, #min node size for splits
) |> 
  set_engine("rpart") |> 
  set_mode("classification")



```


```{r dtree_tune}

#create a model with with tuning specifications
tree_spec_tune <- decision_tree(
  cost_complexity = tune(),
  tree_depth = tune(),
  min_n = tune()
) |> 
  set_engine("rpart") |> 
  set_mode("classification")

#test out some hyperparameters
tree_grid <- grid_regular(cost_complexity(), tree_depth(),
                          min_n(),
                          levels = 5) #try 5 levels for hyperparameters (so 125 total outcomes)

tree_grid



```

```{r dtree_resample}

#use resampling folds (speciifed above) to build trees in parallel

doParallel::registerDoParallel() #build trees in parallel
#200s
tree_rs <- tune_grid(
  tree_spec_tune,
  user~.,
  resamples = cv_folds,
  grid = tree_grid,
  metrics = metric_set(accuracy)
)

tree_rs



```


```{r dtree_metrics}

#table of metrics

#at least one visual of results

```


## Bagged Tree

```{r bagg_model}



```

```{r bagg_resample}



```

```{r bagg_tune}



```

```{r bagg_metrics}

#table of metrics

#at least one visual of results

```

## Random Forest

```{r rf_model}




```

```{r rf_resample}



```

```{r rf_tune}




```

```{r rf_metrics}

#table of metrics

#at least one visual of results


```


## Comparing the models
```{r model_metrics}

#table of metrics

#at least one visual of results

```





